---
title: "lateral_angles"
output: html_document
---

This file is to calculate the angle of fish among upward, forward, and downward swimming based on just the lateral view.

```{r library, message=FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
library(tidyverse)
library(lubridate)
library(signal)
library(patchwork)
library(plotly)
```


## Example of single video data file *DON'T RUN THESE*

```{r}
ms03_20211013_trial01 <- read.csv("/Users/minggong/Documents/Tytell_Lab/Project2/Data/UpAndDownLateral3/videos/20211013_ms03_trial01_lateralDLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.csv")
ms03_20211013_trial02 <- read.csv("/Users/minggong/Documents/Tytell_Lab/Project2/Data/UpAndDownLateral3/videos/20211013_ms03_trial02_lateralDLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.csv")
```

```{r}
combine <- read.csv("/Users/minggong/Documents/Tytell_Lab/Project2/Data/Qualitative/combine.csv")
```

```{r}
dlccolnames <- read_csv("/Users/minggong/Documents/Tytell_Lab/Project2/Data/UpAndDownLateral3/videos/20211013_ms03_trial01_lateralDLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.csv", col_names = FALSE, skip = 1, n_max = 2)

dlccolnames <- paste(dlccolnames[1,], dlccolnames[2,], sep = '_')
dlccolnames[[1]] <- 'frame'

ms03_20211013_trial02 <- read_csv("/Users/minggong/Documents/Tytell_Lab/Project2/Data/UpAndDownLateral3/videos/20211013_ms03_trial02_lateralDLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.csv", skip = 3, col_names = FALSE)
colnames(ms03_20211013_trial02) <- dlccolnames
```

```{r}
lateral_points <-
  ms03_20211013_trial02 %>%
  mutate(Fish = "ms03",
         Date = "10_13_2021",
         Trial = 2)
```

```{r clean dataset}
lateral_points <- tibble(Fish = "ms03",
                         Date = "10_13_2021",
                         Trial = 2,
                         frame = as.integer(ms03_20211013_trial02$scorer),
                         left_eye_x = as.numeric(ms03_20211013_trial02$DLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.24),
                         left_eye_y = as.numeric(ms03_20211013_trial02$DLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.25),
                         left_lateral_strip_x = as.numeric(ms03_20211013_trial02$DLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.33),
                         left_lateral_strip_y = as.numeric(ms03_20211013_trial02$DLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.34),
                         peduncle_middle_x = as.numeric(ms03_20211013_trial02$DLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.18),
                         peduncle_middle_y = as.numeric(ms03_20211013_trial02$DLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.19))

lateral_points <- lateral_points[-(1:2),]

lateral_points <- left_join(x = lateral_points, y = combine, by = c("Fish", "Date", "Trial")) %>%
  select(-c(starting_time, tail_beat, movement_transition, tail_direction, leading)) %>% 
  unique()

```

```{r calculate angles}
#lateral_points <- mutate(lateral_points, slope = (as.numeric(left_lateral_strip_y) - as.numeric(left_eye_y))/(as.numeric(left_lateral_strip_x) - as.numeric(left_eye_x)),
#                         angle = atan2(slope, 1))

lateral_points <- mutate(lateral_points,
                         eye_strip_angle_deg = atan2((left_lateral_strip_y - left_eye_y), 
                                                 (left_lateral_strip_x - left_eye_x)) * 180/pi,
                         strip_peduncle_angle_deg = atan2(peduncle_middle_y - left_lateral_strip_y,
                                                          peduncle_middle_x -left_lateral_strip_x) * 180/pi,
                         eye_peduncle_angle_deg = atan2(peduncle_middle_y - left_eye_y,
                                                        peduncle_middle_x - left_eye_x) * 180/pi)

```

```{r}
ggplot(lateral_points, aes(x=frame)) +
  geom_point(aes(y=eye_strip_angle_deg), color = 'red') +
  geom_point(aes(y=strip_peduncle_angle_deg), color = 'blue') +
  geom_point(aes(y=eye_peduncle_angle_deg))
```





```{r read-in and extracting test}
test_file <- "/Users/minggong/Documents/Tytell_Lab/Project2/Data/UpAndDownLateral3/videos/20211013_ms03_trial02_lateralDLC_resnet50_UpAndDownLateral2Feb7shuffle1_1030000.csv"

str_extract(filename, '20\\d{6}')  # extract 8 numbers in the format of '20XXXXXX'
str_extract(filename, 'ms\\d{2}')  # extract 4 characters & numbers in the format of 'msXX'
str_extract(filename, 'trial\\d+_') %>%
  str_extract('\\d+') %>%
  as.numeric()
```





## All video data files
*START FROM HERE*
```{r customize read-in function}
read_lateral_data <- function(filename) {
  dlccolnames <- read_csv(filename, col_names = FALSE, skip = 1, n_max = 2)
  
  dlccolnames <- paste(dlccolnames[1,], dlccolnames[2,], sep = '_')
  dlccolnames[[1]] <- 'frame'
  
  data <- read_csv(filename, skip = 3, col_names = FALSE)
  colnames(data) <- dlccolnames 
  
  filedate <- str_extract(filename, '20\\d{6}')
  fishid <- str_extract(filename, 'ms\\d{2}')
  trialnum <- str_extract(filename, 'trial\\d+_') %>%
    str_extract('\\d+') %>%
    as.numeric()

  data %>%
    mutate(date = filedate,
           fish = fishid,
           trial = trialnum)
}
```

read_lateral_data(test_file)

```{r}
all_videos <- 
    list.files(pattern = "*.csv")
```

1:12 20211013_ms03
13:29 20211014_ms04
30:42 20211028_ms03
43:56 20211029_ms05
```{r read all data files, message=FALSE, show_col_types = FALSE}
data <-
  all_videos[43:56] %>%
    map_dfr(~read_lateral_data(.)) %>%
  mutate(date = ymd(date))

```


```{r read in trial_info_v1 directly}
# read in trial_info_v1.csv
# It has the starting frame and ending frame (from qualitative data).
trial_info_v1 <- read_csv("/Users/minggong/Documents/Tytell_Lab/Project2/Data/Qualitative/trial_info_v1.csv")

# change date format to year_month_date
trial_info_v1$date <- as.Date(trial_info_v1$date, format = "%m_%d_%Y")
trial_info_v1$date <- ymd(trial_info_v1$date)

# change starting and ending frame into time in second
trial_info_v1$start <- trial_info_v1$start/200
trial_info_v1$end <- trial_info_v1$end/200
```
```{r read in trial_info_v2}
# trial_info_v2 has the range of frames that include both points and left eye and strip

trial_info_v2 <- read_csv("/Users/minggong/Documents/Tytell_Lab/Project2/Data/Qualitative/trial_info_v2.csv")
# change date format to year_month_date
trial_info_v2$date <- as.Date(trial_info_v1$date, format = "%m_%d_%Y")
trial_info_v2$date <- ymd(trial_info_v1$date)

# change starting and ending frame into time in second
trial_info_v2$start_full_range <- trial_info_v2$start_full_range/200
trial_info_v2$end_full_range <- trial_info_v2$end_full_range/200
```

```{r}
# combine all trial information with the big data set
trial_info_v1 <- left_join(trial_info_v1, trial_info_v2, by = c("ID", "fish", "date", "trial", "vertical_direction"))
data <- left_join(data, trial_info_v1, by = c("fish", "date", "trial"))
```


```{r}
# change frame into time in seconds
data <- data %>% 
  mutate(time = frame/200, .after = frame)
```


```{r}
head(data)
tail(data)
```


```{r}
datalong <-
  data %>%
  pivot_longer(cols = caudal_dorsal_top_x:hyoid_likelihood, 
               names_to = c('point', '.value'),
               names_pattern = "(.+)_(x|y|likelihood)")
```

```{r define replace na function}
spline_replace_na <- function(y) {
  good <- !is.na(y)
  x <- seq_along(y)
  
  if (all(good)) {
    y
  } else {
    span <- rep(FALSE, length(y))
    
    a <- which.max(good)
    b <- length(y) + 1 - which.max(rev(good))
    span[a:b] <- TRUE
  
    ys <- rep(NA_real_, length(y))
    sp <- spline(x[good], y[good], xout = x[span])
  
    ys[span] <- sp$y
    ys
  }
}
```
*YOU CAN JUMP TO NEXT "RUN THIS" CHUNK*


```{r define find_jump function}
# find_jump <- function(x, y) {
#   dx = x - lag(x)
#   dy = y - lag(y)
#   
#   summary <- summarise(
#     mean_dx = mean(dx),
#     mean_dy = mean(dy),
#     median_dx = quantile(dx, 0.5),
#     median_dy = quantile(dy, 0.5),
#     threshold_dx = quantile(dx, 0.90, na.rm = T),
#     threshold_dy = quantile(dy, 0.90, na.rm = T))
#   
#   
#   if (dx <= summary$threshold_dx) {
#     x = x
#   } else {
#     x = NA_real_
#   }
#   
#   if (dy <= summary$threshold_dy) {
#     y = y
#   } else {
#     y = NA_real_
#   }
#   
# }
```



```{r}
# Histogram: Plot the distribution of dx and dy to find the best threshold
datalong %>% 
  dplyr::filter(trial == 5) %>% 
  dplyr::filter(likelihood > 0.99) %>% 
  dplyr::select(12:15) %>% 
  group_by(point) %>% 
  mutate(dx = x-lag(x), 
         dy = y-lag(y)) %>%
  plot_ly(x = ~dy, type = "histogram", histnorm = "probability")
```

```{r}
# Boxplot: Plot the distribution of dx and dy to find the best threshold
boxplot_dx_dy <- datalong %>% 
  dplyr::filter(trial == 5) %>% 
  dplyr::filter(likelihood > 0.99) %>% 
  dplyr::select(12:15) %>% 
  group_by(point) %>% 
  mutate(dx = x-lag(x), 
         dy = y-lag(y)) %>%
  plot_ly(x = ~point, y = ~dx, type = "box", name = "dx") %>% 
  add_trace(y = ~dy, type = "box", name = "dy")

print(boxplot_dx_dy)

```

```{r}
# Find threshold number: when likelihood >0.99, what are the mean differences between points?
datalong %>% 
  dplyr::filter(trial == 3) %>% 
  dplyr::filter(likelihood > 0.99) %>% 
  select(12:15) %>% 
  group_by(point) %>% 
  mutate(dx = x-lag(x), 
         dy = y-lag(y)) %>%
  na.omit() %>% 
  summarize(mean_dx_good = mean(dx),
            mean_dy_good = mean(dy))

datalong %>% 
  dplyr::filter(trial == 3) %>% 
  dplyr::filter(likelihood < 0.99) %>% 
  select(12:15) %>% 
  group_by(point) %>% 
  mutate(dx = x-lag(x), 
         dy = y-lag(y)) %>%
  na.omit() %>% 
  summarize(mean_dx = mean(dx),
            mean_dy = mean(dy))

```


```{r}
# test the spline_replace_NA function and find_jump function
datalong %>% 
  dplyr::filter(fish == 'ms03' & trial == 1) %>% 
  dplyr::filter(point == 'left_eye') %>%
  ggplot(aes(x = time)) + 
  geom_line(aes(y = x), color = 'green') + 
  geom_line(aes(y = y), color = 'blue')

datalong %>% 
  dplyr::filter(fish == 'ms03' & trial == 1) %>%
  mutate(x = if_else(likelihood > 0.98, x, NA_real_),
         y = if_else(likelihood > 0.98, y, NA_real_)) %>%
  dplyr::filter(point == 'left_eye') %>%
  mutate(is_jump = find_jump(x, y),
         x = if_else(is_jump, NA_real_, x),
         y = if_else(is_jump, NA_real_, y)) %>%
  mutate(xs = spline_replace_na(x), 
         ys = spline_replace_na(y)) %>%
  ggplot(aes(x = time)) +
  geom_line(aes(y = xs), color = 'red') +
  geom_line(aes(y = x), color = 'green') +
  geom_line(aes(y = ys), color = 'red') +
  geom_line(aes(y = y), color = 'blue')

```


```{r}
## Old chunk of the original codes
# datalong <-
#   datalong %>%
#   mutate(x = if_else(likelihood > 0.98, x, NA_real_),
#          y = if_else(likelihood > 0.98, y, NA_real_)) %>%
#   mutate(is_jump = find_jump(x, y, threshold = 0.5),
#          x = if_else(is_jump, NA_real_, x),
#          y = if_else(is_jump, NA_real_, y)) %>%
#   group_by(point) %>% 
#   rename(x0 = x, y0 = y) %>%
#   mutate(x = spline_replace_na(x0),
#          y = spline_replace_na(y0))
  
```



```{r}
## RUN THIS
# filter points of high likelihood
datalong <- datalong %>% 
  mutate(x = if_else(likelihood > 0.98, x, NA_real_),
         y = if_else(likelihood > 0.98, y, NA_real_)) %>% 
  group_by(ID, point) %>% 
  mutate(dx = x - lag(x),
         dy = y - lag(y)) %>% 
  ungroup()
```

```{r}
## RUN THIS
# calculate threshold for points of high likelihood
thres_summary <- datalong %>% 
  group_by(ID, point) %>% 
  summarise(mean_dx = mean(na.omit(dx)),
            mean_dy = mean(na.omit(dy)),
            median_dx = quantile(dx, 0.5, na.rm = T),
            median_dy = quantile(dy, 0.5, na.rm = T),
            threshold_dx = quantile(dx, 0.90, na.rm = T),
            threshold_dy = quantile(dy, 0.90, na.rm = T))

datalong <- datalong %>% left_join(thres_summary %>% select(ID, point, threshold_dx, threshold_dy), by = c("ID", "point"))
```

```{r}
## RUN THIS
# use the threshold to filter the jump points again
# apply the spline_replace_NA function on the big data set to replace NA points
datalong <- datalong %>% 
  group_by(ID, point) %>% 
  mutate(x = if_else(dx > threshold_dx, NA_real_, x),
         y = if_else(dy > threshold_dy, NA_real_, y)) %>% 
  rename(xs = x, ys = y) %>% 
  mutate(x = spline_replace_na(xs),
         y = spline_replace_na(ys))
```


```{r}
datalong %>% 
  dplyr::filter(ID == 5 & point == 'left_eye') %>%
  ggplot(aes(x = time)) +
    geom_line(aes(y = x), color = 'red') +
    geom_line(aes(y = xs), color = 'green') +
    geom_line(aes(y = y), color = 'red') +
    geom_line(aes(y = ys), color = 'blue') +
    facet_wrap(~ID)
```


```{r}
data <- datalong %>%
  pivot_wider(names_from = 'point', names_glue = '{point}_{.value}', 
              values_from = c(xs, ys, x, y, likelihood, dx, dy, threshold_dx, threshold_dy))
```

```{r subset points}
points <- data %>% select(ID,
                          date, 
                          fish, 
                          trial,
                          vertical_direction,
                          frame,
                          time,
                          start,
                          start_full_range,
                          end,
                          end_full_range,
                          left_eye_x,
                          left_eye_y,
                          spiny_dorsal_ant_base_x,
                          spiny_dorsal_ant_base_y,
                          hyoid_x,
                          hyoid_y,
                          left_lateral_strip_x,
                          left_lateral_strip_y,
                          peduncle_middle_x,
                          peduncle_middle_y)
```


```{r customize function to calculate angle}
angle <- function(x1, y1, x2, y2){
  angle <- atan2((y2 - y1), (x2 - x1)) * 180/pi
  return(angle)
}

# angle(points$left_eye_x, points$left_lateral_strip_x, points$left_eye_y, points$left_lateral_strip_y)  # test
```

```{r}
# calculate angles of the line based on two points
points <- points %>% mutate(eye_strip_angle_deg = angle(left_eye_x, left_eye_y, left_lateral_strip_x, left_lateral_strip_y),
                            strip_peduncle_angle_deg = angle(peduncle_middle_x, peduncle_middle_y, left_lateral_strip_x, left_lateral_strip_y),
                            eye_peduncle_angle_deg = angle(peduncle_middle_x, peduncle_middle_y, left_eye_x, left_eye_y),
                            eye_dorsalant_angle_deg = angle(left_eye_x, left_eye_y, spiny_dorsal_ant_base_x, spiny_dorsal_ant_base_y),
                            eye_hyoid_angle_deg = angle(left_eye_x, left_eye_y, hyoid_x, hyoid_y),
                            dorsalant_hyoid_angle_deg = angle(spiny_dorsal_ant_base_x, spiny_dorsal_ant_base_y, hyoid_x, hyoid_y))

```



```{r}
# plot angle vs frame in one specified trial

color_1 <- c("eye_strip_angle_deg" = "blue",
           "strip_peduncle_angle_deg" = "cyan",
           "eye_peduncle_angle_deg" = "deep sky blue",
           "eye_dorsalant_angle_deg" = "deep pink",
           "eye_hyoid_angle_deg" = "red",
           "dorsalant_hyoid_angle_deg" = "maroon")

points %>% 
  dplyr::filter(ID == 5) %>%
  mutate(strip_peduncle_angle_deg = unwrap(strip_peduncle_angle_deg*pi/180)*180/pi) %>%  
  ggplot(aes(x=time, shape=vertical_direction)) +
  geom_point(aes(y=eye_strip_angle_deg, color ='eye_strip_angle_deg')) +
  geom_point(aes(y=strip_peduncle_angle_deg, color = 'strip_peduncle_angle_deg')) +
  geom_point(aes(y=eye_peduncle_angle_deg, color = 'eye_peduncle_angle_deg')) +
  geom_point(aes(y=eye_dorsalant_angle_deg, color = 'eye_dorsalant_angle_deg')) +
  geom_point(aes(y=eye_hyoid_angle_deg, color = 'eye_hyoid_angle_deg')) +
  geom_point(aes(y=dorsalant_hyoid_angle_deg, color = 'dorsalant_hyoid_angle_deg')) +
  labs(x = "Time (sec)", y = "Angle relative to horizontal plane") +
  scale_color_manual(values = color_1)

```

```{r}
# use eye-strip angle as body angle
body_angle_1 <- points %>% 
  select(ID, date, fish, trial, vertical_direction, frame, time, start, end, eye_strip_angle_deg)
# body_angle_1 <- body_angle_1[!(is.na(body_angle_1$eye_strip_angle_deg)), ]
```


```{r}
# customize the filter
tailbeatfreq <- 1
sampfreq <- 1/0.005
## butterfilt <- signal::butter(9, 4*tailbeatfreq / (0.5*sampfreq), "low")  # filter 1
butterfilt <- signal::butter(9, 2*tailbeatfreq / (0.5*sampfreq), "low")  # filter 2 (smoother than filter 1)

# Deal with the edge effect of the butterfilt
EndEffect <- function(filt,x) {
  x <- x[!is.na(x)]
  signal::filtfilt(filt,c(rev(x),x,rev(x)))[(length(x) + 1):(2 * length(x))]
}
```


```{r}
# smooth body angle & calculate angular velocity
body_angle_1 <- body_angle_1 %>%
  dplyr::filter(!is.na(eye_strip_angle_deg)) %>% 
  group_by(ID) %>% 
  mutate(unwrap_angle = unwrap(eye_strip_angle_deg*pi/180)*180/pi,
         body_angle = unwrap_angle-unwrap_angle[1],  # subtract baseline
         body_angle_smooth = EndEffect(butterfilt, body_angle),
         body_angle_smooth_rad = body_angle_smooth*pi/180,
         ang_vel_rads = (lead(body_angle_smooth_rad) - lag(body_angle_smooth_rad))/(2*0.005))


body_angle_1 <- body_angle_1 %>%
  dplyr::group_by(ID) %>%
  fill(c(ang_vel_rads), .direction = "downup") %>%
  mutate(ang_vel_smooth_rads = EndEffect(butterfilt, ang_vel_rads)) %>% 
  dplyr::ungroup()

```



```{r}
# compare the smooth data and the original data
body_angle_1 %>% 
  dplyr::filter(ID == 5) %>%
  ggplot(aes(x=time, shape=vertical_direction)) +
  geom_point(aes(y=body_angle, color ='body_angle')) +
  geom_point(aes(y=body_angle_smooth, color = 'body_angle_smooth')) +
  labs(x = "Time (sec)", y = "Body angle (deg)")
```

```{r}
body_angle_1 %>% 
  dplyr::filter(ID == 44) %>%
  ggplot(aes(x=time, shape=vertical_direction)) +
  geom_point(aes(y=ang_vel_rads_1, color ='ang_vel_rads_1')) +
  geom_point(aes(y=ang_vel_smooth_rads_1, color = 'ang_vel_smooth_rads_1')) +
  geom_point(aes(y=ang_vel_rads_2, color ='ang_vel_rads_2')) +
  geom_point(aes(y=ang_vel_smooth_rads_2, color = 'ang_vel_smooth_rads_2')) +
  labs(x = "Time (sec)", y = "Angular velocity")
```

```{r}
# plot eye-strip angle as body angle
ggplot(body_angle_1, aes(x = time, y = body_angle_smooth, color = vertical_direction)) +
  geom_point() +
  facet_wrap(~ID, scales = "free") +
  labs(x = "Time (sec)", y = "Body angle (eye-strip) (deg)")

```


```{r}
# plot eye-strip angular velocity
ggplot(body_angle_1, aes(x = time)) +
  geom_point(aes(y = ang_vel_rads, color = 'red')) +
  geom_point(aes(y = ang_vel_smooth_rads, color = 'blue')) +
  facet_wrap(~ID, scales = "free") +
  labs(x = "Time (sec)", y = "Angular velocity (eye-strip) (rad/s)")
```





*Strip-peduncle angle is not as good as the eye-strip angle.*
*SKIP*
```{r}
# use strip-peduncle angle as body angle
body_angle_2 <- points %>% 
  select(ID, date, fish, trial, vertical_direction, frame, strip_peduncle_angle_deg)

body_angle_2 <- body_angle_2 %>%
  group_by(ID) %>% 
  mutate(length = length(frame)/8,
         unwrap_angle = unwrap(strip_peduncle_angle_deg*pi/180)*180/pi,
         mean_angle = mean(unwrap_angle[frame<length/8]),
         body_angle = unwrap_angle-mean(unwrap_angle[frame<length/8]),
         body_angle_smooth = EndEffect(butterfilt, body_angle))

#         angular_speed = (body_angle - lag(body_angle, default = first(body_angle)))/10)
```

```{r}
# plot strip-peduncle angle as body angle
ggplot(body_angle_2, aes(x = frame, y = body_angle_smooth, color = vertical_direction)) +
  geom_point() +
  facet_wrap(~ID, scales = "free") +
  labs(x = "Frame", y = "Body angle (eye-strip)")
```

```{r}
# plot strip-peduncle angular speed
ggplot(body_angle_2, aes(x = frame, y = angular_speed, color = vertical_direction)) +
  geom_point() +
  facet_wrap(~ID, scales = "free") +
  labs(x = "Frame", y = "Angular velocity (eye-strip) (deg/s)")
```




*START FROM HERE AGAIN*
```{r horizontal and vertical speed}
# calculate horizontal speed and vertical speed based on the peduncle middle point
# unit should be BL/s

speed <- points

speed <- points %>% 
  group_by(ID) %>%
  mutate(fish_length = mean(abs(left_eye_x - peduncle_middle_x), na.rm = TRUE),
         left_eye_vx = (lead(left_eye_x) - lag(left_eye_x))/(2*0.005),
         left_eye_vx_BL = left_eye_vx/fish_length,
         left_eye_vy = (lead(left_eye_y) - lag(left_eye_y))/(2*0.005),
         left_eye_vy_BL = left_eye_vy/fish_length)# %>% 
#  fill(c(left_eye_vx_BL, left_eye_vy_BL), .direction = "downup")

#speed <- speed[!is.na(speed$left_eye_vx_BLs),]
#speed <- speed[!is.na(speed$left_eye_vy_BLs),]

```


```{r}
# Deal with the edge effect of the butterfilt
EndEffectDF <- function(df, filt, var) {
  x <- df %>%
    pull({{var}})
  df %>%
    mutate("{{var}}_s" := signal::filtfilt(filt,c(rev(x),x,rev(x)))[(length(x) + 1):(2 * length(x))],
           "{{var}}_s2" := signal::filtfilt(filt,x))
}
```

```{r}
speed <- 
  speed %>%
  dplyr::filter(!is.na(left_eye_vy_BL)) %>%
  group_by(ID) %>%
  group_modify(~ EndEffectDF(.x, butterfilt, left_eye_vy_BL)) %>%
  ungroup()

  # group_by(ID) %>%
  # nest() %>% 
  # mutate(smoothed = purrr::map(data, ~ EndEffect(butterfilt, .x$left_eye_vy_BLs)))

```

```{r}
speed <- 
  speed %>%
  dplyr::filter(!is.na(left_eye_vx_BL)) %>%
  group_by(ID) %>%
  group_modify(~ EndEffectDF(.x, butterfilt, left_eye_vx_BL)) %>%
  ungroup()
```

```{r}
speed %>%
  ggplot(aes(x = frame)) +
  geom_line(aes(y = left_eye_vy_BL, color = 'left_eye_vy_BL')) +
  geom_line(aes(y = left_eye_vy_BL_s, color = 'left_eye_vy_BL_s')) +
  geom_line(aes(y = left_eye_vy_BL_s2, color = 'left_eye_vy_BL_s2')) +
  facet_wrap(~ ID, scales = 'free')
```

```{r}
speed %>%
  ggplot(aes(x = frame)) +
  geom_line(aes(y = left_eye_vx_BL, color = 'left_eye_vx_BL')) +
  geom_line(aes(y = left_eye_vx_BL_s, color = 'left_eye_vx_BL_s')) +
  geom_line(aes(y = left_eye_vx_BL_s2, color = 'left_eye_vx_BL_s2')) +
  facet_wrap(~ ID, scales = 'free')
```

```{r}
# plot left eye x position
points %>% 
  dplyr::filter(vertical_direction == "U") %>% 
  ggplot(aes(shape = vertical_direction)) +
  geom_point(aes(x = time, y = left_eye_x, color = "left_eye_x")) +
  facet_wrap(~ID, scales = "free") +
  labs(x = "Time (sec)", y = "Left Eye x Position")
```


```{r}
# plot horizontal-forward velocity
speed %>% 
  dplyr::filter(vertical_direction == "D") %>%   
  ggplot(aes(shape = vertical_direction)) +
  geom_point(aes(x = time, y = left_eye_vx_BLs)) +
  facet_wrap(~ID, scales = "free") +
  labs(x = "Time (sec)", y = bquote('Speed'~(BL*s^-1)))

```

```{r}
# ggsave("speed_up.png", plot = last_plot(), width = 2560, height = 1440, units = "px", limitsize = FALSE)
```



*Plot the body angle, speed, tail 3D position for an example trial*
```{r}
# pick one trial as an example
ID3 <- data %>% 
  dplyr::filter(fish == 'ms03' & trial == 3) %>%

```


```{r}
ID3_long <-
  ID3 %>%
  pivot_longer(c(contains('eye'), contains('angle')), names_to = "name", values_to = "value")
```


```{r}
ggplot(ID3_long, aes(x = time, y = value, color = name)) +
  geom_point() +
  facet_wrap(~name, ncol = 1, scales = "free_y") +
  scale_color_discrete() +
  theme(legend.position = "none")
```


```{r}
# 1. body angle
p1 <- body_angle_1 %>% 
  dplyr::filter(ID == 3) %>%
  ggplot(aes(x=time, y=body_angle_smooth, color = 'green')) +
  geom_point() +
  labs(x = "Time (sec)", y = "Body angle (deg)") +
  theme(legend.position = "none")
```


```{r}
# 2. left_eye_vx_BLs
ID3<-na.omit(ID3)
ID3$left_eye_vx_BLs<-signal::filtfilt(butterfilt, ID3$left_eye_vx_BLs)

p2 <- ggplot(ID3, aes(x = time, y = left_eye_vx_BLs)) +
  geom_point() +
  labs(y = "x velocity (BL/s)")
```


```{r}
# 3. left_eye_vy_BLs
ID3$left_eye_vy_BLs <- signal::filtfilt(butterfilt, ID3$left_eye_vy_BLs)

p3 <- ggplot(ID3, aes(x = time, y = left_eye_vy_BLs)) +
  geom_point() +
  labs(y = "y velocity (BL/s)")
```   



```{r}
# read in previous 3D model data
ID3_3d <- read_csv("/Users/minggong/Documents/Tytell_Lab/Project2/Code/ID3_3d.csv")

ID3_3d <- ID3_3d %>% 
  mutate(time = frame/200, .after=frame) %>% 
  pivot_longer(cols = 3:7, names_to = "bodypart", values_to = "value")

# reorder factors
ID3_3d$bodypart <- factor(ID3_3d$bodypart, 
                             levels = c("dorsal_top_p_top", "dorsal_middle_p_top", "middle", "ventral_middle_p_bottom", "ventral_bottom_p_bottom"),
                             labels = c("dorsal top", "dorsal middle", "middle", "ventral middle", "ventral bottom"))
```


```{r}
p4 <- ggplot(ID3_3d, aes(x = time, y = value, color = bodypart)) +
  geom_point() + 
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        legend.text = element_text(size=8)) +
  labs(y = "Amplitude")
p4
```


```{r}
# use plotly to choose the traces you want
plot_ly(data = ID3_3d, x = ~time) %>%
  add_markers(y = ~dorsal_top_p_top, name = "dorsal_top_p_top") %>%
  add_markers(y = ~dorsal_middle_p_top, name = "dorsal_middle_p_top") %>% 
  add_markers(y = ~middle, name = "middle") %>% 
  add_markers(y = ~ventral_middle_p_bottom, name = "ventral_middle_p_bottom") %>% 
  add_markers(y = ~ventral_bottom_p_bottom, name = "ventral_bottom_p_bottom")
```


```{r}
# use patchwork to combine pictures
# notice that patchwork can only combine ggplot2 plots

left_eye_trial3 <- p1/p2/p3/p4
left_eye_trial3
```




*Combine the data and output a csv with all results (body angle, angular velocity, vertical velocity).*
```{r}
lateral_angle <- body_angle_1 %>% 
  ungroup() %>%  
  select(ID,
         date,
         fish,
         trial,
         frame,
         time,
         body_angle_smooth,
         body_angle_smooth_rad,
         ang_vel_smooth_rads)

speed <- speed %>%
  select(ID,
         date,
         fish,
         trial,
         frame,
         left_eye_vy_BL_s,
         left_eye_vx_BL_s)

lateral_angle <- left_join(lateral_angle, speed, by = c("ID", "date", "fish", "trial", "frame"))
```

```{r}
write_csv(lateral_angle, "lateral_angle_20211029_ms05.csv")
```

