---
title: "quantitative_analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggsignif)
```

```{r read and arrange data}
quan_v4 <- read.csv("quantitative_v4.csv")
quan_v4 <- select(quan_v4, 1:10) %>% na.omit()
quan_v4 <- quan_v4 %>% 
  group_by(fish, date, trial) %>% 
  mutate(id = cur_group_id()) %>% 
  ungroup() %>% 
  relocate(id, .before = fish)
str(quan_v4)
```

```{r distribution}
ggplot(data = quan_v4, mapping = aes(x = frame, y = position, color = lobe, shape = fish)) +
  geom_point()
# position centered around -0.5

ggplot(data = quan_v4, mapping = aes(x = frame, y = position, color = vertical_direction)) +
  geom_point()

ggplot(data = quan_v4, mapping = aes(x = vertical_direction, y = position, color = lobe)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.5) # no difference

ggplot(data = quan_v4, mapping = aes(x = as.factor(tail_beat), y = position, color = lobe)) +
  geom_boxplot() # no difference
```

## Section added by Eric

Check whether position differs for lobe, depending on direction and tail beat number
```{r}
ggplot(data = quan_v4, mapping = aes(x = as.factor(tail_beat), y = position, color = lobe)) +
  geom_boxplot() +
  facet_grid(vertical_direction ~ .)
```

Here we take the mean position for each lobe in each tail beat for each vertical direction,
then pivot the data wider so that we have the mean position for each lobe in different columns.
Then we can take the difference and look at that for each tail beat.
```{r}
quan_v4 %>%
  group_by(tail_beat, vertical_direction, lobe) %>%
  summarize(meanpos = mean(position, na.rm = TRUE)) %>%
  pivot_wider(values_from = c(meanpos), names_from = lobe) %>%
  mutate(asym = dorsal - ventral) %>%
  ggplot(aes(x = tail_beat, y = asym, color = vertical_direction)) +
  geom_point()
```
Looks like there is something going on! Although the asymmetry increases over tail beats, which is weird. There may be some baseline asymmetry, which may be why the horizontal points hover around 0.2 - 0.4.

Now let's look at this for all of the tail beats.

First, set up a new data frame with columns that correspond to dorsal vs ventral lobe, for position, phase_shift and frame. Then we can take the difference.
```{r}
asym_data <-
  quan_v4 %>%
  pivot_wider(names_from = lobe, values_from = c(position, phase_shift, frame)) %>%
  mutate(asym = position_dorsal - position_ventral) 

asym_summary <- asym_data %>% 
  group_by(vertical_direction) %>% 
  summarize(mean = mean(asym), sd = sd(asym), max = max(asym), min = min(asym))
```

This plots the mean +- 95% confidence limit for the asymmetry.
```{r}
asym_data %>%
  ggplot(aes(x = tail_beat, y = asym, color = vertical_direction)) +
  stat_summary(fun.data = 'mean_cl_boot', position = position_dodge(width = 0.3)) +
  labs(x = "Half tail beat", y = "Amplitude asymmetry", color = "Vertical direction") + 
  scale_colour_manual(labels = c("Downward", "Horizontal", "Upward"), values = c("#fe988d","#0b5394", "#a5cfb5")) + # change color label
  scale_x_continuous(breaks = c(1,2,3,4,5,6)) + 
  theme_bw()

  geom_signif(y_position = c(0.68, 0.68, 0.68, 0.68), 
              xmin = c(2.9, 3.9, 4.9, 5.9), 
              xmax = c(3.1, 4.1, 5.1, 6.1), 
              annotation = c("**", "**", "*", "**"), 
              tip_length = 0, size = 0.5) +
  geom_signif(y_position = c(0.62, 0.62, 0.62, 0.62), 
              xmin = c(2.9, 3.9, 4.9, 5.9), 
              xmax = c(3.1, 4.1, 5.1, 6.1), 
              annotation = c("**", "*", "*", "*"), 
              tip_length = 0, size = 0.5) +
  geom_signif(y_position = c(0.45), 
              xmin = c(4), 
              xmax = c(4.1), 
              annotation = c("*"), 
              tip_length = c(0.01, 0.15),
              size = 0.5) +
  geom_signif(y_position = c(0.32), 
              xmin = c(6), 
              xmax = c(6.1), 
              annotation = c("*"), 
              tip_length = c(0.01, 0.1),
              size = 0.5)

```
Still looks like there's a difference!

```{r tail beat 1}
asym_anova_1 <- aov(asym ~ vertical_direction, 
                  data = filter(asym_data, tail_beat == 1))
summary(asym_anova_1)
TukeyHSD(asym_anova_1)
```

```{r tail beat 2}
asym_anova_2 <- aov(asym ~ vertical_direction, 
                  data = filter(asym_data, tail_beat == 2))
summary(asym_anova_2)
TukeyHSD(asym_anova_2)
```


Simple ANOVA shows that it's significant in tail beat 3.
```{r tail beat 3}
asym_anova_3 <- aov(asym ~ vertical_direction, 
                  data = filter(asym_data, tail_beat == 3))
summary(asym_anova_3)
TukeyHSD(asym_anova_3)
```
END OF ERIC's ADDITIONS.

```{r tail beat 4}
asym_anova_4 <- aov(asym ~ vertical_direction, 
                  data = filter(asym_data, tail_beat == 4))
summary(asym_anova_4)
TukeyHSD(asym_anova_4)
```

```{r tail beat 5}
asym_anova_5 <- aov(asym ~ vertical_direction, 
                  data = filter(asym_data, tail_beat == 5))
summary(asym_anova_5)
TukeyHSD(asym_anova_5)
```

```{r tail beat 6}
asym_anova_6 <- aov(asym ~ vertical_direction, 
                  data = filter(asym_data, tail_beat == 4))
summary(asym_anova_6)
TukeyHSD(asym_anova_6)
```


```{r plot phase shift}
# (playback) frame rate of mp4 video == 10 fps
# (recording) frame rate of cine video == 200 fps
# quan_v2 <- mutate(quan_v2, phase_shift_sec = phase_shift/200) # calculate phase shift in second

quan_v4 <- quan_v4 %>%
  group_by(id) %>% 
  mutate(# tailbeat_frame = max(frame)-min(frame), # frame range of 3 tail beats
    tailbeat_frame = (max(frame) - min(frame))/3, # I would use this if it's 3 tail beats
         ratio = phase_shift/tailbeat_frame) %>%    # then this is the phase shift as a fraction of the tail beat cycle
  ungroup()

# when phase shift < 0, dorsal-leading; when phase shift > 0, ventral-leading

phase_shift_summary <- quan_v4 %>% 
  na.omit() %>% 
  group_by(vertical_direction) %>% 
  summarize(number = length(ratio), mean = mean(ratio), sd = sd(ratio), max = max(ratio), min = min(ratio))

ggplot(data = quan_v4, mapping = aes(x = vertical_direction, y = ratio)) + 
  geom_boxplot() +
  labs(x = "Vertical direction", y = "Ratio of phase shift") +
  scale_x_discrete(labels = c("Downward", "Horizontal", "Upward")) +
  theme_bw() +
  geom_signif(comparisons = list(c("D", "U")), map_signif_level = T, y = 0.15) +
  geom_signif(comparisons = list(c("D", "H")), map_signif_level = T, y = 0.13)

```

```{r compare phase shift}
two_way_anova <- aov(ratio ~ vertical_direction, data = quan_v4)
summary(two_way_anova)
TukeyHSD(two_way_anova)
```

