---
title: "R Notebook"
output: word_document
---

```{r setup}
library(tidyverse)
library(plotly)
library(here)
library(ggpmisc)
library(pracma)
```

```{r}
source('fin_processing_3d.R')
```

```{r read data}
filename <- here('20211028_ms03_trial08_DLC_3D.csv')
```

```{r}
colnames <- read_csv(filename, skip=1, n_max=2, col_names = FALSE)
```

The first three rows of the DLC output file contain
* The scorer, which isn't important for us
* The body part names
* The coordinate name (x, y, z)

Below, we read in the 2nd and 3rd rows, and then merge them so that we get, for example 'dorsal_top_x'.
```{r}
bodyparts <- as.character(colnames[1,])
coord <- as.character(colnames[2,])

colnames <- tibble(bodyparts = bodyparts[2:length(bodyparts)], coord = coord[2:length(coord)])
colnames <-
  colnames %>%
  mutate(colname = str_c(bodyparts, '_', coord)) %>%
  pull(colname)

colnames
```

Now read in the rest of the file with good column names
```{r}
data <- read_csv(filename, skip=3, col_names = c('frame', colnames))

head(data)
```

```{r}
bodypartorder = c("dorsal_top", "dorsal_middle", "middle", "ventral_middle", "ventral_bottom", "peduncle_top", "peduncle_middle", "peduncle_bottom", "eye")
```

This plots just one point in 3D
```{r}
plot_ly(data, x = ~middle_x, y = ~middle_y, z = ~middle_z) %>%
  add_markers()
```

This makes our data set "tidy", arranging all of the x, y, z values in one column each, with another column that tells us which body part we're looking at.
```{r}
data_long <-
  data %>% 
  pivot_longer(cols = -frame, names_to = c('bodypart', '.value'), names_pattern = '(.+)_([xyz])')

head(data_long)
```

```{r}
data_long <-
  data_long %>%
  mutate(bodypart = factor(bodypart, levels = bodypartorder))
```

This is all of the points!
```{r}
plot_ly(data_long, x = ~x, y = ~y, z = ~z, color = ~bodypart) %>%
  add_markers()
```

It looks like decreasing y is up, decreasing x is forward, and decreasing z is left. Let's make new coordinates that are called `up`, `fwd`, and `left`.
```{r}
data_long <-
  data_long %>%
  mutate(up = median(y, na.rm = TRUE) - y, fwd = max(x, na.rm = TRUE) - x, left = median(z, na.rm = TRUE) - z)

head(data_long)
```

Now that seems to be oriented correctly!
```{r}
plot_ly(data_long, x = ~fwd, y = ~left, z = ~up, color = ~bodypart) %>%
  add_markers()
```

```{r}
triangles <- tribble(
  ~i, ~j, ~k,
  'dorsal_top',    'dorsal_middle', 'peduncle_top',
  'dorsal_middle',  'middle', 'peduncle_middle',
  #'peduncle_top', 'dorsal_middle', 'peduncle_middle',
  'ventral_middle', 'peduncle_middle', 'middle',
  'ventral_bottom', 'peduncle_bottom', 'ventral_middle')
triangles
```

```{r}
tri.ind <-
  triangles %>%
  mutate(across(c(i,j,k), ~ factor(.x, levels = bodypartorder)),
         across(c(i,j,k), ~ as.integer(.x) - 1),
         label = i+1)

tri.ind
```

```{r}
data_long %>%
  arrange(frame, bodypart) %>%
  filter(bodypart != 'eye') %>%
  filter(frame >= 60 & frame < 80) %>%
  plot_ly(x = ~fwd, y = ~left, z = ~up,
          i = tri.ind$i, j = tri.ind$j, k = tri.ind$k, frame = ~frame,
          intensity = tri.ind$label, intensitymode='cell',
          type = 'mesh3d') %>%
  add_markers(color = ~bodypart)
```


```{r}
data_long %>%
  plot_ly(x = ~frame, y = ~left, color = ~bodypart) %>%
  add_markers()
```


```{r}
p1 <-
  data_long %>%
  filter(bodypart %in% c('ventral_bottom')) %>% # c('dorsal_top', 'ventral_bottom', 'peduncle_top', 'peduncle_middle', 'peduncle_bottom')) %>%
  # smooth_and_interp(k = 5) %>%
  plot_ly(x = ~frame, y = ~left, color = ~bodypart) %>%
  add_markers()

p2 <-
  data_long %>%
  filter(bodypart %in% c('ventral_bottom')) %>% # c('dorsal_top', 'ventral_bottom', 'peduncle_top', 'peduncle_middle', 'peduncle_bottom')) %>%
  # smooth_and_interp(k = 5) %>%
  plot_ly(x = ~frame, y = ~up, color = ~bodypart) %>%
  add_markers()

subplot(p1, p2, nrows = 2, shareX = TRUE)
```

```{r}
data_nojumps <-
  data_long %>%
  group_by(bodypart) %>%
  group_modify(~find_jump(.x, frame, fwd, left, up, bigjump = 0.99)) %>%
  mutate(left = if_else(isjump == 1, NA_real_, left)) %>%
  ungroup() %>%
  smooth_and_interp(frame, fwd,left,up, k = 1)
```

```{r}
data_nojumps %>%
  ggplot(aes(x = frame, color = bodypart)) +
  geom_point(aes(y = left)) +
  geom_line(aes(y = left_s))
```

```{r if points cannot be connected}
require(tidyverse)

connect_points_in_frame <- function(df, connect,
                           coords = c('fwd', 'left', 'up'))
{
  df2 <- data.frame(matrix(vector(), 0, 2 + 2*length(coords)))
  colnames(df2) <- c('from', 'to', str_c(coords, '1'), str_c(coords, '2'))
  
  for (i in seq(1, nrow(connect))) {
    a <- connect$from[i]
    b <- connect$to[i]
    df2[i, 'from'] <- a
    df2[i, 'to'] <- b
    
    for (c in coords) {
      ac <- paste(a, c, sep = '_')
      bc <- paste(b, c, sep = '_')
      
      c1 <- paste0(c, '1')
      c2 <- paste0(c, '2')
      
      df2[i, c1] <- df[1, ac]
      df2[i, c2] <- df[1, bc]
    }    
  }
  df2
}

connect_points <- function(df, connect,
                           coords = c('fwd', 'left', 'up'))
{
  df <-
    df %>%
    select(frame, bodypart, one_of(coords))
  
  df <- df %>% pivot_wider(names_from = bodypart, values_from = one_of(coords), names_glue = "{bodypart}_{.value}")
  
  df %>%
    group_by(frame) %>%
    group_modify(~ connect_points_in_frame(.x, connect))
}

smooth_and_interp_point <- function(t, x, k = 7)
{
  xs <- rep_along(x, NA)
  
  good <- !is.na(x)
  i <- which(good)
  a <- i[1]
  b <- i[length(i)]
  
  if (k == 1) {
    xm <- x
  }
  else {
    xm <- runmed(x, k, endrule = 'median', na.action = 'na.omit')
  }
  
  sp <- smooth.spline(t[good], xm[good])
  val <- predict(sp, t[a:b])
  
  xs[a:b] <- val$y
  xs
}

smooth_and_interp <- function(df, t, x, y, z, 
                              k = 7)
{
  df %>%
    group_by(bodypart) %>%
    mutate("{{x}}_s" := smooth_and_interp_point({{t}}, {{x}}, k = k),
           "{{y}}_s" := smooth_and_interp_point({{t}}, {{y}}, k = k),
           "{{z}}_s" := smooth_and_interp_point({{t}}, {{z}}, k = k))
}

find_jump <- function(df, t, x, y, z,
                      bigjump = 0.98,
                      maxjumpdur = 10,
                      key = NA)
{
  if (!is.na(key)) {
    print(str(key))
  }
  
  df <- df %>%
    mutate(dx = ({{x}} - lag({{x}})) / ({{t}} - lag({{t}})),
           dy = ({{y}} - lag({{y}})) / ({{t}} - lag({{t}})),
           dz = ({{z}} - lag({{z}})) / ({{t}} - lag({{t}})),
           jumpmag = sqrt(dx^2 + dy^2 + dz^2))

  qx <- quantile(df$jumpmag, bigjump, na.rm = TRUE)
  
  matchjumps <-
    df %>%
    filter(jumpmag > qx) %>%
    mutate(jumpdir = (dx*lag(dx) + dy*lag(dy) + dz*lag(dz)) / (jumpmag * lag(jumpmag)),
           jumpsize = jumpmag / lag(jumpmag)) %>%
    mutate(isjumpback = (jumpdir < -0.95) & (jumpmag/lag(jumpmag) - 1 < 0.1),
           jumpdur = ({{t}}) - lag({{t}}))
  
  matchjumps$jumpdir <- 0
  
  if (nrow(matchjumps) < 2) {
    df %>%
      mutate(jumpdir = 0,
             isjump = FALSE)
  } else {
    matchedprev <- FALSE
    for (i in seq(2,nrow(matchjumps))) {
      if (!matchedprev & matchjumps$isjumpback[i] & (matchjumps$jumpdur[i] < maxjumpdur)) {
        matchjumps$jumpdir[i-1] <- 1
        matchjumps$jumpdir[i] <- -1
        matchedprev <- TRUE
      }
      else {
        matchedprev <- FALSE
      }
    }
    
    matchjumps <-
      matchjumps %>%
      select({{t}}, jumpdir, isjumpback)
    
    df %>%
      left_join(matchjumps, by = as_label(enquo(t))) %>%
      mutate(jumpdir = replace_na(jumpdir, 0),
             isjump = cumsum(jumpdir))
  }
}
```

```{r plot connected points}
data_nojumps %>%
  group_by(frame) %>%
  summarize(dorsal_lobe = left_s[1] - left_s[6],
         ventral_lobe = left_s[5] - left_s[8]) %>%
  ungroup()  %>%
  plot_ly(x = ~frame) %>%
  add_markers(y = ~dorsal_lobe, name = "Dorsal") %>%
  add_markers(y = ~ventral_lobe, name = "Ventral")
```

```{r find peaks}
data_nojumps_plot <- data_nojumps %>%
  group_by(frame) %>%
  summarize(dorsal_lobe = left_s[1] - left_s[6],
         ventral_lobe = left_s[5] - left_s[8]) %>%
  ungroup()

data_peaks <- data_frame(lobe = "dorsal",
                     frame = data_nojumps_plot$frame[ggpmisc:::find_peaks(data_nojumps_plot$dorsal_lobe)], 
                     position = data_nojumps_plot$dorsal_lobe[ggpmisc:::find_peaks(data_nojumps_plot$dorsal_lobe)]) %>%
  add_row(lobe = "dorsal",
          frame = data_nojumps_plot$frame[ggpmisc:::find_peaks(-data_nojumps_plot$dorsal_lobe)],
          position = data_nojumps_plot$dorsal_lobe[ggpmisc:::find_peaks(-data_nojumps_plot$dorsal_lobe)]) %>%
  add_row(lobe = "ventral",
          frame = data_nojumps_plot$frame[ggpmisc:::find_peaks(data_nojumps_plot$ventral_lobe)],
          position = data_nojumps_plot$ventral_lobe[ggpmisc:::find_peaks(data_nojumps_plot$ventral_lobe)]) %>% 
  add_row(lobe = "ventral",
          frame = data_nojumps_plot$frame[ggpmisc:::find_peaks(-data_nojumps_plot$ventral_lobe)],
          position = data_nojumps_plot$ventral_lobe[ggpmisc:::find_peaks(-data_nojumps_plot$ventral_lobe)]) 
data_peaks <- data_peaks[order(data_peaks$frame),] # sort all rows by the frame number
data_peaks_1 <- data_peaks %>% 
  subset(frame >= 230 & frame <= 600) # create a subset by matching the range of frame numbers with qualitative analysis
data_peaks_1 <- data_peaks_1[-c(),] # Remove certain rows if needed. The final data frame should have 12 rows, which means 3 tail beats.

# write_csv(data_peaks_1, "3d_20211013_ms03_trial03.csv")

```

```{r}
str(data_nojumps_plot)
findpeaks(data_nojumps_plot$dorsal_lobe, minpeakdistance = 20, sortstr = F)
```

```{r}
# obtain the phase shift

data_nojumps_phase <- data_nojumps %>%
  group_by(frame) %>%
  summarize(dorsal_lobe = left_s[1] - left_s[6],
         ventral_lobe = left_s[5] - left_s[8]) %>%
  ungroup() %>% 
  slice(231:601)

ccf(data_nojumps_phase$dorsal_lobe, data_nojumps_phase$ventral_lobe, lag = 80, pl = F)

max(abs(ccf(data_nojumps_phase$dorsal_lobe, data_nojumps_phase$ventral_lobe, lag = 100, plot = TRUE)[["acf"]]))

lagindex <- which.max(abs(ccf(data_nojumps_phase$dorsal_lobe, data_nojumps_phase$ventral_lobe, lag = 100, plot = FALSE)[["acf"]]))
bestlag <- ccf(data_nojumps_phase$dorsal_lobe, data_nojumps_phase$ventral_lobe, lag =100, plot = FALSE)[["lag"]][[lagindex]] %>% print()
```


```{r}
# output a csv file of five points on the tail
data_nojumps_plot <- data_nojumps %>% 
  select(bodypart, frame, left_s) %>% 
  group_by(frame) %>% 
  summarize(dorsal_top_p_top = left_s[1] - left_s[6],
            dorsal_middle_p_top = left_s[2] - left_s[6],
            middle = left_s[3] -left_s[7],
            ventral_middle_p_bottom = left_s[4] - left_s[8],
            ventral_bottom_p_bottom = left_s[5] - left_s[8]) %>%
  ungroup()

ifelse(file.exists("20211028_ms03_trial08_3dtail.csv"), yes = askYesNo("Do you want to continue? Check file name!"), no = write_csv(data_nojumps_plot, "20211028_ms03_trial08_3dtail.csv"))
```

